---
title: "A Noob Linear Mixed Effects Model Example"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

There are a few packages we can use, but the most popular is lme4. There's one caveat with this, and that is the fact it doesn't print out p-values. Given that we hedge everything on our trusty, arbitrary alpha level, we need to load in another package that will give us the info we need.

```{r}
library(lme4)
library(lmerTest)
```

Instead of using AgeingIQ, I'm using some kind of made up data from another one of my studies. The variables are:

1. Participant ID
2. Working Memory Load (1,3 or 5 letters displayed)
3. Reaction time (RT) (s)
4. Accuracy (Acc) (%)
5. Physical Activity (High/Low)

The repeated measure in this data set are RT and Acc for each WM load (within subjects factor). Physical Activity is a between subjects factor. 

What we are interested in determining in this data set is whether there is an interaction between working memory performance at increasing difficulty and physical activity. There are a few ways we can tackle this. You could do a repeated measures ANOVA. However, the whole point of this tutorial is to run a mixed effects model which allows us to control for our random effects (in this case, participant ID). 

Let's just look at the main effects. In order to view the results of the model, the anova() function is used.

```{r}

model <- lmer(Acc ~ WM_Load + PhysicalActivity + (1|Participant), data = lmeData)

anova(model)

```

To add an interaction term, you change the + to a *:

```{r}

modelInt <- lmer(Acc ~ WM_Load*PhysicalActivity + (1|Participant), data = lmeData)

anova(modelInt)

```
